---
title: "Capitulo 1: Time Series Data"
output: html_notebook
---
1.4 Plots, trends and seasonal variation 

```{r}
AP<-AirPassengers
plot(AirPassengers,ylab="Passengers (thousands)")
```
```{r}
class(AP);
start(AP);
end(AP);
frequency(AP)
```
A função aggregate permite visualizar a tendencia melhor, ao remover a sazonalidade da série.
A função cycle permite quebrar a série nas seasons
```{r}
layout(1:2);
plot(aggregate(AP));
boxplot(AP~cycle(AP))
```
#Base de dados do IBGE
```{r}
desemprego<-DesempBrasil
desemp_total<-ts(desemprego[,7],start =2012, frequency=4 )
desemp_total.annual.ts<-aggregate(desemp_total)/4
layout(1:2)
plot(desemp_total.annual.ts)
plot(desemp_total)
```
```{r}
#library(fpp2)
layout(1:2)
ggseasonplot(desemp_total, polar=TRUE)
ggseasonplot(desemp_total, polar=FALSE)
```
```{r}
Desemp.Q1<-window(desemp_total, start=c(2012,1), freq=TRUE)
Desemp.Q2<-window(desemp_total, start=c(2012,2), freq=TRUE)
Desemp.Q3<-window(desemp_total, start=c(2012,3), freq=TRUE)
Desemp.Q4<-window(desemp_total, start=c(2012,4), freq=TRUE)
Q1.ratio<-mean(Desemp.Q1)/mean(desemp_total)
Q2.ratio<-mean(Desemp.Q2)/mean(desemp_total)
Q3.ratio<-mean(Desemp.Q3)/mean(desemp_total)
Q4.ratio<-mean(Desemp.Q4)/mean(desemp_total)
ratio.quart<-c(Q1.ratio,Q2.ratio,Q3.ratio,Q4.ratio)
ratio.quart

```
#US Unemployement Rate
```{r}
www<-"http://www.maths.adelaide.edu.au/andrew.metcalfe/Data/USunemp.dat"
US.month<-read.table(www, header=T)
attach(US.month)
US.month.ts<-ts(USun,start=c(1996,1),end=c(2006,10),freq=12)
plot(US.month.ts, ylab="% Unemployed Rate")
```
#1.4.3 Multiple Time Series
```{r}
www<-"http://www.maths.adelaide.edu.au/andrew.metcalfe/Data/cbe.dat"
CBE<-read.table(www,header=T)
#CBE[1:4,]
#class(CBE)
Elec.ts<-ts(CBE[,3],start=1958,freq=12)
Beer.ts<-ts(CBE[,2],start = 1958,freq=12)
Choc.ts<-ts(CBE[,1],start = 1958,freq=12)
plot(cbind(Elec.ts,Beer.ts,Choc.ts))
    
```
```{r}
AP.elec<-ts.intersect(AP,Elec.ts)
start(AP.elec)
end(AP.elec)
AP.elec[1:3,]

AP<-AP.elec[,1]
Elec<-AP.elec[,2]
layout(1:2)
plot(AP, ylab="AirPassengers in thousands")
plot(Elec, ylab="Electricity production in MkWh")
plot(as.vector(AP),as.vector(Elec),xlab="AirPassengers",ylab="ElectricityProduction")
abline(reg=lm(Elec~AP))
```
#1.4.4 Quarterly Exchange Rate: GBP to NZ dollar
```{r}
www<-"http://www.maths.adelaide.edu.au/andrew.metcalfe/Data/pounds_nz.dat"
Z<-read.table(www,header=T)
Z[1:4,]
Z.ts<-ts(Z, st=1991,fr=4)
plot(Z.ts, xlab="Years",ylab="Quarterl Exchange Rate in $NZ / pound")
```
```{r}
z.92.96<-window(Z.ts,st=c(1992,1),end=c(1996,1))
z.96.98<-window(Z.ts,st=c(1996,1),end=c(1998,1))
layout(1:2)
plot(z.92.96,xlab="Years",ylab="Quarterl Exchange Rate in $NZ / pound")
plot(z.96.98,xlab="Years",ylab="Quarterl Exchange Rate in $NZ / pound")
```
#1.4.5 Global Temperatura series
```{r}
www<-'http://www.maths.adelaide.edu.au/andrew.metcalfe/Data/global.dat'
#Global<-scan(www)
Global.ts<-ts(Global,st=c(1856,1),end=c(2005,12),fr=12)
layout(1:2)
plot(Global.ts)
Global.annual<-aggregate(Global.ts,FUN=mean)
plot(Global.annual)
```
```{r}
New.series<-window(Global.ts,start=c(1970,1),end=c(2005,12))
New.time<-time(New.series)
plot(New.series)
abline(reg=lm(New.series~New.time))
```
#1.5.5 Decomposition in R
```{r}
Elec.decom<-decompose(Elec.ts)
plot(Elec.decom)
```
```{r}
Trend<-Elec.decom$trend
Seasonal<-Elec.decom$seasonal
ts.plot(cbind(Trend,Trend*Seasonal),lty=1:2)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
